Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-07-04 src/tex/ms.tex
[showyourwork] overleaf sync

% Author list
\author[0000-0002-6299-7542]{Ekaterina Ilin }
\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}

@@

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. These flares are not expected to appear different from intrinsic stellar flares individually. However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage obtained from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Since 2018, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of both missions for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

@@

%Short intro to SaB mechanism

%Short intro to AW mechanism

% ----------------------------------------------------------------------------------

% In both models, the radius of the perturbing body, and its proximity increase the power of interaction. Would it not be more reasonable to search for star-brown dwarf or star-star interactions?

% Flares caused by magnetic interactions in close binary systems were previously observed. Flares detected on the young eccentric binary system V 773 Tau could be attributed to colliding magnetospheres~\citep{massi2002periodic, massi2008interacting}, while other promising systems, like DQ Tau, another young equal-mass binary system in a strongly eccentric orbit, did not live up to the expectations~\citep{salter2008captured,salter2010recurring,getman2011young,getman2016search, kospal2018spots}. In these systems, it is, however, unclear whether it is mass transfer or reconnection that drive excess coronal emission and flares at periastron. Substellar objects can orbit the star much closer than a stellar companion without filling the Roche lobe. Gravitationally, a rocky planet can go unnoticed by its host even if it orbits at ultrashort periods of less than a day. 
% Interactions in systems with brown dwarf companions are less likely than binary systems

@@

%We note that we found K2-354 under EPIC 211314705 and K2-329(!) in \cite{bouma2020cluster}, who refer to the detection paper \cite{pope2016transiting}, and also under TIC 468989066.

@@

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flaring. Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm bandpass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

@@

Among the stars without given uncertainties, the rotation of GJ 3082 was measured and is consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ 674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has rotation periods inferred from activity indicators~\citep{astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674, and GJ 3323.
\end{itemize}

@@

The convective turnover time $\tau$ in  $R \rm o=P_{rot}/\tau$ is derived using bolometric luminosity~(Section~\ref{sec:data:lum}), following~\citet{reiners2014generalized, reiners2022magnetism}:

@@

\citet{fischer2019timevariable} argue that the synodic period, at which the sub-planetary point crosses the same (magnetic) surface element on the star, may result in a clearer interaction signal, because a magnetically active region that is prone to planet-induced flares, will be passed only once per that period. However, we repeated the coherence calculation for $P_{rot}$, and found that the ratio is $>0.1$ for most stars~(Figure~\ref{fig:coherence_hist}, top panel), so that a statistical analysis of rotational flare modulation is not feasible. %We argue, however, that in a transiting system with high orbital inclination, the effect with synodic period reduces to modulation with the orbital period:

%If the orbital axis inclination is high, then the passage will regularly occur on the back of the star relative to the observer. The true period will therefore be additionally modulated by the visibility, resulting in an intermittency of the signal. As a consequence, to detect a modulation with the synodic period, a significantly longer observing baseline is needed than to detect a modulation with the orbital period. Additionally, if there are many active regions (at the same time, or in sequence) that produce flares when the sub-planetary footpoint passes across, the effect with the synodic period further converges to a modulation with orbital period. 
%In our sample, the observing baseline is multiple years long for many systems. Most stars, even the most rigidly rotating late M dwarfs, show variability in their spot modulation on such long timescales~\citep[e.g.,][]{giles2017kepler,davenport2015spots, namekata2019lifetimes}).

% Even if differential rotation was absent and active regions were extremely stable, we would likely be unable to observe the passage of the sub-planetary point over the flare-prone region every synodic period.

%However, this assumes that the landscape of active regions on the star is stable throughout the entire observing time, and that the number of such regions in the footpoint passageway is low enough to produce a periodic modulation. Note that this is a different effect from the geometrical modulation of visibility we look for in modulation with the orbital period in this study.

%Even for the few stars, for which the rotational period is known precisely enough, we need to make numerous assumptions. These include the number of active regions crossed by the footpoint of interaction, differential rotation, spot evolution, and a significant correlation between the location of the planet and the location of the planet-induced flare on the stellar surface. For modulation with the orbital period, we require only that last correlation between footpoint location and orbital phase.

%There are three reasons for this: First, the uncertainty for the rotation was not always provided, in which case we assigned a $10\%$ error. Second, rotation period is more difficult to ascertain from a light curves' variability than the orbital period, particularly if the modulation is weak and/or rapidly evolving. And third, differential rotation causes spots at different latitudes to move at different speeds, so it becomes difficult to even define a rotation period, as long as we don't know at which latitude(s) the flares occur~(see also Section~\ref{sec:discussion:rotsyn}).

@@

For instance, Kepler-808 (KOI-1300) has a $\sim0.4M_\odot$ \citep{kraus2016impact} companion that is about 1.8 mag fainter than the primary at a separation of $0.78\arcsec$~\citep{baranec2016roboao}. TOI-837 has co-moving M dwarf a few arcsec away in the same young cluster, 5 mag fainter, but the primary is an F9-G0 star, so flares could still originate from both stars~\citep{bouma2020cluster}. HD 41004 B, DS Tuc A, and LTT 1445 are known multiple systems where all components contribute significantly to the total TESS flux.

@@

\paragraph{KOI-12}
\label{sec:results:individualstars:koi12}
KOI-12, also known as Kepler-448, is a two-planet system with an eccentric outer planet KOI-12 c, and an inner Warm Jupiter planet KOI-12 b in a $17.8\,$d orbit~\citep{masuda2017eccentric}. With respect to rotation rate, stellar and planetary radius, the KOI-12 system is similar to HIP~67522 and TAP~26 (although for the those young stars the radii are large because the stars are still contracting onto the main sequence). In constrast, KOI-12 is $1.4\pm0.3\,$Gyr old~\citep{bourrier2015sophie}, an F5 sub-giant~\citep{frasca2016activity}, which is already evolving off of the main sequence. Yet, despite its high expected power of interaction, we detect no excess flaring in phase with KOI-12 b. The reason could be that KOI-12 b is in fact super-Alfv\'enic due to the relatively wide orbit compared to TAP 26 and HIP 67522. Our estimate of the Alfv\'en radius place KOI-12 b in the sub-Alfv\'enic regime~(see Fig.~\ref{fig:as}), but the scaling law applied there may be inaccurate for post-main sequence stars. Intermittent interaction might play a role here as well, and could be favored if future observations capture episodes during which the interaction is 'on'.

@@

Magnetic and tidal SPI are difficult to disentangle by their global effects on stellar activity indicators~(Section~\ref{sec:intro:global}). Locally, however, we can discriminate between the two by their relevant periods, that is, $P_{\rm orb}$ and $P_{\rm orb /2}$. In Section~\ref{sec:discussion:tidal}, we repeat the analysis in this paper with $P_{orb}/2$, and find yet more tentative trends in line with scaling laws for tidal interaction. In a system-by-system comparison, we find consistency with the expectation that slowly rotating, likely old, systems can only show local tidal interaction, while rapidly rotating systems with strong magnetic fields are dominated by magnetic forces.
\subsection{Alfv\'en surface}
\label{sec:discussion:as}

@@

\subsection{Eccentricity}
\label{sec:discussion:eccentricity}
Orbital eccentricity can affect magnetic SPI by modulating the planet's distance to the star. If the planetary orbit is highly eccentric, the difference between periastron and apoastron can be significantly larger than the variability of the Alfv\'en radius. If then apoastron is outside this radius, and periastron within, the planet may find itself in the sub-Alfv\'enic zone only briefly, during the planet's rapid periastron passage. This was previously exploited in the case of HD 17156~\citep{maggio2015coordinated}, and analogously for the colliding magnetospheres in binary systems~\citep{massi2002periodic,getman2016search,das2023discovery}. In our sample, the estimates for eccentricities of the innermost planets are all moderate to low, with $e\leq 0.35$. The exception is K2-25, with a relatively high $e\approx0.43$. If periastron passage in K2-25 occurs when the planet is in front of the star relative to the observer, the visibility effect of the interaction footpoint is further increased by the narrow phase range of the periastron passage. \citet{stefansson2020habitable} estimated the argument of periastron $\omega=120^{+12}_{-14}\,$deg for K2-25 (with longitude of the ascending node chosen such that $\omega$ is the same as the longitude of periastron, with $\omega=90\,$deg being periastron passage during transit, see~\citealt{kipping2010investigations,dawson2012photoeccentric}). We can therefore conjecture that the tentative signal of flaring SPI in K2-25 is enhanced by eccentricity. Statistically, we are more likely to observe a planetary transit close to periastron, since periastron permits more orbital inclinations to transit than apoastron. Overall, we expect eccentric transiting systems to experience elevated SPI signal, which could be tested in the future when better constraints on eccentricity become available~(see Table~\ref{tab:maintable_lit}).

@@

\subsection{Intermittency}
\label{sec:discussion:intermittency}
Magnetic SPI power in a star-planet system may sometimes be low for episodes of time, and high for others. HD~179949 is the prototype example for this phenomenon. The orbital modulation of its chromospheric indicators was observable for four out of six epochs on a five-year baseline~\citep{shkolnik2003evidence,shkolnik2008nature}. It is not clear how long a continuous epoch of interaction or non-interaction might be. This leaves room for many different explanations. Uninterrupted, long-term monitoring of individual systems is required to investigate the reasons for the temporary cessation of interaction. Candidate mechanisms for the star include: 
\begin{itemize}
    \item the Alfv\'en radius moving in- and out the planet's orbit, either over the course of an activity cycle, or through short-term variations in the magnetic field and wind properties caused by, e.g., coronal mass ejections;
    \item the interacting footpoint moving to higher latitudes due to changes in the large scale field, which reduces the modulation of its visibility 
    \item an active latitude (i.e., small scale field) moving away from the magnetic footpoints passageway; or
    \item the available magnetic energy decreasing such that the flares or other magnetic interaction indicators still occur but fall below the detection threshold for a period of time.
    \item the availability of a sufficient amount of plasma in the vicinity of the magnetic interaction region to yield significant observable flare emission in the stellar atmosphere (in analogy to the Jupiter-Io system, wherein Io's volcanism provides the plasma that eventually produces an interaction hotspot on Jupiter,~\citealt{clarke1996farultraviolet, prange1996rapid})
\end{itemize}

@@

%Interpreting non-detections as effects of viewing geometry in fully aligned systems ignores this caveat.

%If we assume that the magnetic dipole axis and the rotation axis of the star are relatively well-aligned, the spin-orbit alignment may have an influence on interaction strength. For instance,  HAT-P-11 rotates nearly pole-on, and HAT-P-11 b is transiting in a near-polar orbit~\citep{bourrier2023dream}.

% \subsection{(Mis-)alignment between orbital, rotational and stellar magnetic axis}
% \label{sec:discussion:misalignment}

%  Relaxing the assumption of alignment between rotational and magnetic dipole axis, we introduce a further complication that is not only due to viewing geometry.

% or more broadly system architecture including misalignment of the magnetic axis. See HAT-P-11, or instance, and the many polar orbits in \cite{bourrier2023dream}

% \cite{miskovetz2022resolving}  found background stars that were not co-moving HAT-P-11.

% HD 179949 has a strong misalignment of about 70 deg between rotational and magnetic axis. Time-resolved X-ray observations~\citep{acharya2022xray} suggest that the magnetic interaction footpoint is close to the stellar rotational pole, which would be close to the magnetic equator of the strongly dipolar field of HD 179949

@@

No observations of local tidal SPI exist yet. However, in the extreme case of RS CVn binaries, starspots often appear modulated with half the orbital period of the system~\citep{olah2002time, ozavci2018recurrent, kriskovics2023ei}. A similar effect was measured when the companion was likely a very low mass star or a brown dwarf~\citep{donati1995activity, frasca2008spots, parks2021interferometric}, and in the case of a white dwarf companion~\citep{hussain2006spot, watson2007roche}. In ellipsoidal binaries, a modulation of flares with $P_{\rm orb}/2$ beyond geometrical effects was observed with Kepler~\citep{gao2016whitelight}, suggesting flaring regions at the location of the extreme tidal bulges in these stars. However, all these systems are synchronized, $P_{rot}=P_{\rm orb}$, so that intrinsic active longitudes~\citep{usoskin2007longterm, weber2013theory, jarvinen2005spots, lanza2009corot} cannot be disambiguated from the result of tidal star-star interaction ~\citep{holzwarth2003dynamics}.

@@

The first model measures the torque $\partial L_{\rm conv} / \partial t$, exerted by angular momentum transfer between star and planet, assuming the star has a convective envelope~(\citealt{penev2012constraining}, their Eqns. 1 and 2). We choose the tidal quality factor $Q_*=10^7$, i.e., low efficiency of tidal dissipation, as a conservative estimate. Our results in Fig~\ref{fig:tidal} (left panel) do not depend on the choice of $Q_*$, as we only need to know the tidal torque down to a proportionality factor. We also make the simplifying assumption that the rotation period of the convective envelope is the same as the surface rotation period measured as described in Section~\ref{sec:data:rotationperiods}, ignoring differential rotation effects.

The second model considers the tidal dissipation timescale -- the shorter the timescale, the higher the power of the interaction~\citep{zahn1977tidal}. Following~\citet{ilic2022tidal}, we adopt Eqn. 2 in~\citet{albrecht2012obliquities} for the tidal dissipation timescale in the convective envelope. We assume that all stars in our sample have a convective envelope because they flare, and flaring requires a dynamo-generated magnetic field, which can only operate in stars with convective envelopes or fully convective stars. Most of our stars are dwarf stars below the Kraft break at spectral type F5~\citep{kraft1967studies}, except for KOI-12, which is an F5 sub-giant~\citep{frasca2016activity}.

The third model uses the gravitational perturbation as a proxy for the interaction~\citep{cuntz2000stellar}, originally derived for SPI. Their Eqn. 1 proposes a gravitational perturbation proportional to the mass ratio between planet and star, and $a^{-3}$.

\subsubsection{Planetary and stellar masses}
\label{sec:discussion:mpms}
We collect the planetary and stellar masses from the literature~(Table~\ref{tab:tidal}). For missing planetary masses, the radii were known, so we used the mass-radius relations implemented in \texttt{astro-forecaster}~\citep[Ben Cassesse's implementation of \texttt{forecaster}][]{chen2017probabilistic}. GJ 3323, GJ 674 and GJ 3082 were also missing stellar mass estimates, which we obtained from the \citep{mann2015how, mann2016erratum} relations between absolute Ks magnitude~\citep{skrutskie2006two} and stellar mass for M dwarfs using the distances from \citet{bailer-jones2021estimating}. For Kepler-42 b, we used  $M_p=0.1-2.06M_\oplus$ that covers pure rock to pure iron compositions in \citet{muirhead2012characterizing}, and adopted the logarithmic mean for our estimate. Kepler-1558 c has a similar radius to Kepler-42 b, i.e., most likely a bare core planet, so we used density ranging from Earth bulk density to pure rock to estimate its mass. We also adopt the upper limit of $5M_J$ for HIP 67522 from~\citet{rizzuto2020tess}, but use the lower error from \citet{chen2017probabilistic}. The assumptions we made here are conservative, and the estimates based on \citet{chen2017probabilistic} use a broad sample of planets, so we consider the planetary mass estimates with their respective uncertainties robust. We note, however, that the mass estimates for RV-detected planets are $M_p \sin i$, so that their expected interaction may be stronger than indicated in Fig.~\ref{fig:tidal}.

@@

Overall, tidal interaction, if at all present, is less pronounced than magnetic SPI in our sample. However, if the observed deviations from intrinsic flaring do in fact represent low level tidal interaction signal, we can conclude that close-in planets can interact both magnetically and tidally. In line with theoretical considerations~\citep{strugarek2017fate}, our results tentatively suggest that magnetic interaction is more pronounced for planets with fast rotating, active hosts, while tidal interactions affect both systems with slowly and rapidly rotating stars. Finally, exceptions such as TAP 26 exemplify that the simple scaling laws we applied here may not capture the complexity of the interaction.

@@

We conducted the so far largest search for flaring star-planet interactions (SPI). Using over 7200 Quarters and Sectors of photometric monitoring from Kepler and TESS archives, we searched for flares in over 1800 systems of the about 3000 listed in the NASA Exoplanet Archive. When flares could be unambiguously attributed to the planet host, and showed three or more energetic events in the data, we searched for flares that clustered in orbital phase. Our final sample consisted of 25 systems, among them well-known flaring hosts like Proxima Cen, AU Mic, and TRAPPIST-1.

Applying both the stretch-and-break and Alfv\'en wing mechanisms for magnetic interaction, we found a tentative trend between the expected power of interaction, and the presence of excess flaring in phase with the orbital period $P_{\rm orb}$. In particular, we found that there may be two branches at high expected power -- one where the measured orbital clustering increases with power, and one without a trend. While it is unlikely that the extent of the Alfv\'en surface is a reason, intermittency in the interaction and the viewing geometry of the systems could explain the branching. Among the investigated systems, the flares in HIP 67522, a young Sun-like Hot Jupiter host, cluster most distinctly, consistent with its deep embedding in the sub-Alfv\'enic zone and its high expected power of interaction.

We also searched the same data for signs of locally expressed tidal interaction, that is, excess flaring in phase with the tidal bulges ($P_{\rm orb}/2$). We found a similar trend with different scaling laws of tidal interaction, albeit at even lower significance. Our system-by-system comparison suggests that young systems, like HIP 67522, may be dominated by magnetic interaction, whereas old, magnetically inactive systems, like Kepler-42, might mostly interact tidally.

>>>>

% Author list
\author[0000-0002-6299-7542]{Ekaterina Ilin}
\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}

@@

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. These flares are not expected to appear different from intrinsic stellar flares individually. However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{\rm orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage obtained from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been cataloged~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Since 2018, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of both missions for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

@@

@@

@@

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flaring. Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm band pass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

@@

Among the stars without given uncertainties, the rotation of GJ~3082 was measured and is consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ~674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ~3323 only has rotation periods inferred from activity indicators~\citep{astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima~Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P_{\rm rot}/2$, as with GJ~3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ~674, and GJ~3323.
\end{itemize}

@@

The convective turnover time $\tau$ in  $R \rm o=P_{\rm rot}/\tau$ is derived using bolometric luminosity~(Section~\ref{sec:data:lum}), following~\citet{reiners2014generalized, reiners2022magnetism}:

@@

\citet{fischer2019timevariable} argue that the synodic period, at which the sub-planetary point crosses the same (magnetic) surface element on the star, may result in a clearer interaction signal, because a magnetically active region that is prone to planet-induced flares, will be passed only once per that period. However, we repeated the coherence calculation for $P_{\rm rot}$, and found that the ratio is $>0.1$ for most stars~(Figure~\ref{fig:coherence_hist}, top panel), so that a statistical analysis of rotational flare modulation is not feasible.

@@

For instance, Kepler-808 (KOI-1300) has a $\sim0.4M_\odot$ \citep{kraus2016impact} companion that is about 1.8 mag fainter than the primary at a separation of $0.78\arcsec$~\citep{baranec2016roboao}. TOI-837 has co-moving M dwarf a few arcsec away in the same young cluster, 5 mag fainter, but the primary is an F9-G0 star, so flares could still originate from both stars~\citep{bouma2020cluster}. HD~41004~B, DS~Tuc~A, and LTT~1445 are known multiple systems where all components contribute significantly to the total TESS flux.

@@

\paragraph{KOI-12}
\label{sec:results:individualstars:koi12}
KOI-12, also known as Kepler-448, is a two-planet system with an eccentric outer planet KOI-12 c, and an inner Warm Jupiter planet KOI-12 b in a $17.8\,$d orbit~\citep{masuda2017eccentric}. With respect to rotation rate, stellar and planetary radius, the KOI-12 system is similar to HIP~67522 and TAP~26 (although for the those young stars the radii are large because the stars are still contracting onto the main sequence). In contrast, KOI-12 is $1.4\pm0.3\,$Gyr old~\citep{bourrier2015sophie}, an F5 sub-giant~\citep{frasca2016activity}, which is already evolving off of the main sequence. Yet, despite its high expected power of interaction, we detect no excess flaring in phase with KOI-12 b. The reason could be that KOI-12 b is in fact super-Alfv\'enic due to the relatively wide orbit compared to TAP 26 and HIP 67522. Our estimate of the Alfv\'en radius place KOI-12 b in the sub-Alfv\'enic regime~(see Fig.~\ref{fig:as}), but the scaling law applied there may be inaccurate for post-main sequence stars. Intermittent interaction might play a role here as well, and could be favored if future observations capture episodes during which the interaction is 'on'.

@@

Magnetic and tidal SPI are difficult to disentangle by their global effects on stellar activity indicators~(Section~\ref{sec:intro:global}). Locally, however, we can discriminate between the two by their relevant periods, that is, $P_{\rm orb}$ and $P_{\rm orb /2}$. In Section~\ref{sec:discussion:tidal}, we repeat the analysis in this paper with $P_{\rm orb}/2$, and find yet more tentative trends in line with scaling laws for tidal interaction. In a system-by-system comparison, we find consistency with the expectation that slowly rotating, likely old, systems can only show local tidal interaction, while rapidly rotating systems with strong magnetic fields are dominated by magnetic forces.
\subsection{Alfv\'en surface}
\label{sec:discussion:as}

@@

\subsection{Eccentricity}
\label{sec:discussion:eccentricity}
Orbital eccentricity can affect magnetic SPI by modulating the planet's distance to the star. If the planetary orbit is highly eccentric, the difference between periastron and apoastron can be significantly larger than the variability of the Alfv\'en radius. If then apoastron is outside this radius, and periastron within, the planet may find itself in the sub-Alfv\'enic zone only briefly, during the planet's rapid periastron passage. This was previously exploited in the case of HD~17156~\citep{maggio2015coordinated}, and analogously for the colliding magnetospheres in binary systems~\citep{massi2002periodic,getman2016search,das2023discovery}. In our sample, the estimates for eccentricities of the innermost planets are all moderate to low, with $e\leq 0.35$. The exception is K2-25, with a relatively high $e\approx0.43$. If periastron passage in K2-25 occurs when the planet is in front of the star relative to the observer, the visibility effect of the interaction footpoint is further increased by the narrow phase range of the periastron passage. \citet{stefansson2020habitable} estimated the argument of periastron $\omega=120^{+12}_{-14}\,$deg for K2-25 (with longitude of the ascending node chosen such that $\omega$ is the same as the longitude of periastron, with $\omega=90\,$deg being periastron passage during transit, see~\citealt{kipping2010investigations,dawson2012photoeccentric}). We can therefore conjecture that the tentative signal of flaring SPI in K2-25 is enhanced by eccentricity. Statistically, we are more likely to observe a planetary transit close to periastron, since periastron permits more orbital inclinations to transit than apoastron. Overall, we expect eccentric transiting systems to experience elevated SPI signal, which could be tested in the future when better constraints on eccentricity become available~(see Table~\ref{tab:maintable_lit}).

@@

\subsection{Intermittency}
\label{sec:discussion:intermittency}
Magnetic SPI power in a star-planet system may sometimes be low for episodes of time, and high for others. HD~179949 is the prototype example for this phenomenon. The orbital modulation of its chromospheric indicators was observable for four out of six epochs on a five-year baseline~\citep{shkolnik2003evidence,shkolnik2008nature}. It is not clear how long a continuous epoch of interaction or non-interaction might be. This leaves room for many different explanations. Candidate mechanisms for the star include: 
\begin{itemize}
    \item the Alfv\'en radius moving in- and out the planet's orbit, either over the course of an activity cycle, or through short-term variations in the magnetic field and wind properties caused by, e.g., coronal mass ejections;
    \item the interacting footpoint moving to higher latitudes due to changes in the large scale field, which reduces the modulation of its visibility; 
    \item an active latitude (i.e., small scale field) moving away from the magnetic footpoints passageway;
    \item the available magnetic energy decreasing such that the flares or other magnetic interaction indicators still occur but fall below the detection threshold for a period of time; or
    \item the availability of a sufficient amount of plasma in the vicinity of the magnetic interaction region to yield significant observable flare emission in the stellar atmosphere (in analogy to the Jupiter-Io system, wherein Io's volcanism provides the plasma that eventually produces an interaction hotspot on Jupiter,~\citealt{clarke1996farultraviolet, prange1996rapid})
\end{itemize}

@@

@@

No observations of local tidal SPI exist yet. However, in the extreme case of RS CVn binaries, starspots often appear modulated with half the orbital period of the system~\citep{olah2002time, ozavci2018recurrent, kriskovics2023ei}. A similar effect was measured when the companion was likely a very low mass star or a brown dwarf~\citep{donati1995activity, frasca2008spots, parks2021interferometric}, and in the case of a white dwarf companion~\citep{hussain2006spot, watson2007roche}. In ellipsoidal binaries, a modulation of flares with $P_{\rm orb}/2$ beyond geometrical effects was observed with Kepler~\citep{gao2016whitelight}, suggesting flaring regions at the location of the extreme tidal bulges in these stars. However, all these systems are synchronized, $P_{\rm rot}=P_{\rm orb}$, so that intrinsic active longitudes~\citep{usoskin2007longterm, weber2013theory, jarvinen2005spots, lanza2009corot} cannot be disambiguated from the result of tidal star-star interaction ~\citep{holzwarth2003dynamics}.

@@

The first model measures the torque $\partial L_{\rm conv} / \partial t$, exerted by angular momentum transfer between star and planet, assuming the star has a convective envelope~(\citealt{penev2012constraining}, their Eqns.~1~and~2). We choose the tidal quality factor $Q_*=10^7$, i.e., low efficiency of tidal dissipation, as a conservative estimate. Our results in Fig~\ref{fig:tidal} (left panel) do not depend on the choice of $Q_*$, as we only need to know the tidal torque down to a proportionality factor. We also make the simplifying assumption that the rotation period of the convective envelope is the same as the surface rotation period measured as described in Section~\ref{sec:data:rotationperiods}, ignoring differential rotation effects.

The second model considers the tidal dissipation timescale -- the shorter the timescale, the higher the power of the interaction~\citep{zahn1977tidal}. Following~\citet{ilic2022tidal}, we adopt Eq.~2 in~\citet{albrecht2012obliquities} for the tidal dissipation timescale in the convective envelope. We assume that all stars in our sample have a convective envelope because they flare, and flaring requires a dynamo-generated magnetic field, which can only operate in stars with convective envelopes or fully convective stars. Most of our stars are dwarf stars below the Kraft break at spectral type F5~\citep{kraft1967studies}, except for KOI-12, which is an F5 sub-giant~\citep{frasca2016activity}.

The third model uses the gravitational perturbation as a proxy for the interaction~\citep{cuntz2000stellar}, originally derived for SPI. Their Eq.~1 proposes a gravitational perturbation proportional to the mass ratio between planet and star, and $a^{-3}$.

\subsubsection{Planetary and stellar masses}
\label{sec:discussion:mpms}
We collect the planetary and stellar masses from the literature~(Table~\ref{tab:tidal}). For missing planetary masses, the radii were known, so we used the mass-radius relations implemented in \texttt{astro-forecaster}~\citep[Ben Cassesse's implementation of \texttt{forecaster},][]{chen2017probabilistic}. GJ 3323, GJ 674 and GJ 3082 were also missing stellar mass estimates, which we obtained from the \citep{mann2015how, mann2016erratum} relations between absolute $K_s$ magnitude~\citep{skrutskie2006two} and stellar mass for M dwarfs using the distances from \citet{bailer-jones2021estimating}. For Kepler-42 b, we used  $M_p=0.1-2.06M_\oplus$ that covers pure rock to pure iron compositions in \citet{muirhead2012characterizing}, and adopted the logarithmic mean for our estimate. Kepler-1558 c has a similar radius to Kepler-42 b, i.e., most likely a bare core planet, so we used density ranging from Earth bulk density to pure rock to estimate its mass. We also adopt the upper limit of $5M_J$ for HIP 67522 from~\citet{rizzuto2020tess}, but use the lower error from \citet{chen2017probabilistic}. The assumptions we made here are conservative, and the estimates based on \citet{chen2017probabilistic} use a broad sample of planets, so we consider the planetary mass estimates with their respective uncertainties robust. We note, however, that the mass estimates for RV-detected planets are $M_p \sin i$, so that their expected interaction may be stronger than indicated in Fig.~\ref{fig:tidal}.

@@

Overall, tidal interaction, if at all present, is less pronounced than magnetic SPI in our sample. However, if the observed deviations from intrinsic flaring do in fact represent low level tidal interaction signal, we can conclude that close-in planets can interact both magnetically and tidally. In line with theoretical considerations~\citep{strugarek2017fate}, our results tentatively suggest that magnetic interaction is more pronounced for planets with fast rotating, active hosts, while tidal interactions dominate systems with slowly rotating, inactive stars. Finally, exceptions such as TAP 26 exemplify that the simple scaling laws we applied here may not capture the complexity of the interaction.

@@

We conducted the so far largest search for flaring star-planet interactions (SPI). Using over 7200 Quarters and Sectors of photometric monitoring from Kepler and TESS archives, we searched for flares in over 1800 systems of the about 3000 listed in the NASA Exoplanet Archive. When flares could be unambiguously attributed to the planet host, and showed three or more energetic events in the data, we searched for flares that clustered in orbital phase. Our final sample consisted of 25 systems, among them well-known flaring hosts like Proxima~Cen, AU~Mic, and TRAPPIST-1.

Applying both the stretch-and-break and Alfv\'en wing mechanisms for magnetic star-planet interaction, we found a tentative trend between the expected power of interaction, and the presence of excess flaring in phase with the orbital period $P_{\rm orb}$. In particular, we found that there may be two branches at high expected power -- one where the measured orbital clustering increases with power, and one without a trend. While it is unlikely that the extent of the Alfv\'en surface is a reason, intermittency in the interaction and the viewing geometry of the systems could explain the branching. Among the investigated systems, the flares in HIP~67522, a young Sun-like Hot Jupiter host, cluster most distinctly in orbital phase, consistent with its deep embedding in the sub-Alfv\'enic zone and its high expected power of interaction.

We also searched the same data for signs of locally expressed tidal interaction, that is, excess flaring in phase with the tidal bulges ($P_{\rm orb}/2$). We found a similar trend with different scaling laws of tidal interaction, albeit at even lower significance. Our system-by-system comparison suggests that young systems, like HIP~67522, may be dominated by magnetic interaction, whereas old, magnetically inactive systems, like Kepler-42, might mostly interact tidally.
